---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chk

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/poissonconsulting/chk.svg?branch=master)](https://travis-ci.com/poissonconsulting/chk)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/chk?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/chk)
[![Codecov test coverage](https://codecov.io/gh/poissonconsulting/chk/branch/master/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/chk?branch=master)
[![License: CC0](https://img.shields.io/badge/License-CC0-blue.svg)](https://creativecommons.org/publicdomain/zero/1.0/)
<!-- badges: end -->

`chk` is an R package for developers to check user-supplied function arguments.

It is designed to be simple, customisable and fast.

## Installation

To install the latest release version from [CRAN](https://cran.r-project.org)
```r
install.packages("chk")
```

To install the latest development version from [GitHub](https://github.com/poissonconsulting/chk)
```r
# install.packages("remotes")
remotes::install_github("poissonconsulting/chk")
```

## Demonstration

### Simple

`chk` provides simple commonly used checks as (`chk_` functions) which can be combined together for more complex checking.

```{r, error = TRUE}
library(chk)

y <- "a"

chk_flag(y)
chk_string(y)

data <- data.frame(x = 1:2)
chk_is(data, "data.frame")
chk_range(nrow(data), c(3,8))
chk_match(nrow(data), c(3,8))

chk_identical(data$x, 2:1)

z <- "b"
chkor(chk_flag(z), chk_number(z))
```

By default, each check returns TRUE if successful or throws an informative error.

Error messages follow the [tidyverse style guide](https://style.tidyverse.org/error-messages.html).

### Customisable

The functions are designed to be customisable.

#### Custom Error Messages

If `err = FALSE` a check returns FALSE (instead of throwing an error) on check failure.

```{r}
chk_flag(1, err = FALSE)
```

This allows developers to provide their own error messages.

```{r, error = TRUE}
if(!chk_flag(1, err = FALSE)) stop("x MUST be a flag (try as.logical())")
```

#### Copy and Paste

`chk` is released under the [CC0](https://creativecommons.org/publicdomain/zero/1.0/) licence.
This allows developers to copy and paste functions into their package
which is useful if they want to minimise their dependencies.

Copied functions should not be exported (to avoid namespace conflicts) and the 
following creditation preserved.
```r
#  Licence: CC
#  Repository: https://github.com/poissonconsulting/chk
```

### Fast

The functions are designed to be fast.

#### Check First

As exemplified by `chk_flag`, the `chk_` functions immediately evaluate their 
object in one line of code and return TRUE if the check is successful.
```{r}
chk_flag
```

#### Minimal Checking

As they are not expected to be directly exposed to users the `chk_` functions 
don't check any of their arguments (other than the object of interest of course!).

#### Turn Off Checking

If a function is being called internally the checks can be turned off as follows
```{r, error = TRUE}
fun <- function(x, ...) {
  if(is_chk_on()) {
    chk_flag(x)
    chk_unused(...)
  }
  x
}

wrapper_on_fun <- function(x) {
  if(is_chk_on()) {
    chk_off()
    on.exit(chk_on())
  }
  fun(x)
}

fun(FALSE) # calls fun with checking as being called by user
wrapper_on_fun(FALSE) # calls fun without checking as being used internally
```

It is only worth doing this if the checks are substantially slower
than the time required to test and turn checking on and off (see [Benchmarking chk]( https://poissonconsulting.github.io/chk/articles/benchmarking-chk.html))

## Inspiration

- [datacheckr](https://github.com/poissonconsulting/datacheckr/)
- [checkr](https://github.com/poissonconsulting/checkr/)
- [err](https://github.com/poissonconsulting/err/)
- [testthat](https://github.com/r-lib/testthat/)

## Contribution

Please report any [issues](https://github.com/poissonconsulting/chk/issues).

[Pull requests](https://github.com/poissonconsulting/chk/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/poissonconsulting/chk/blob/master/CODE_OF_CONDUCT.md).
By contributing, you agree to abide by its terms.
