---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chk

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/poissonconsulting/chk.svg?branch=master)](https://travis-ci.com/poissonconsulting/chk)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/chk?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/chk)
[![Codecov test coverage](https://codecov.io/gh/poissonconsulting/chk/branch/master/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/chk?branch=master)
[![License: CC0](https://img.shields.io/badge/License-CC0-blue.svg)](https://creativecommons.org/publicdomain/zero/1.0/)
<!-- badges: end -->

`chk` is an R package for developers to check user-supplied inputs to their functions.

It is designed to be simple, fast and customisable.

## Installation

To install the latest release version from [CRAN](https://cran.r-project.org)
```r
install.packages("chk")
```

To install the latest development version from [GitHub](https://github.com/poissonconsulting/chk)
```r
# install.packages("remotes")
remotes::install_github("poissonconsulting/chk")
```

## Demonstration

### Simple

`chk` provides simple commonly used checks which can be combined together
for more complex checking.

```{r, error = TRUE}
library(chk)

y <- "a"

chk_flag(y)
chk_string(y)

chkor(chk_flag(y), chk_number(y))
chkor(chk_flag(y), chk_string(y))
```

As the functions are not expected to be used in pipes they simply return a flag (non-missing logical scalar).

### Fast

The functions are designed to be fast. 

As they are not expected to be exposed to users the `chk` functions don't check any of their inputs (other than the first argument of course!).

Checking user supplied inputs is vital but it slows code down when those checks are unnecessary ie the code is called internally.
To allow conditional checking the developer can use something like

```{r, error = TRUE}
fun <- function(x, chk = TRUE) {
  if(isTRUE(chk)) {
    chk_flag(x)
  }
  x
}

fun(1) # when called by user
fun(1, chk = FALSE) # when called internally
```

### Customisable

The functions are designed to be customisable.

If `err = FALSE` the `chk` functions return FALSE on check failure.
```{r}
chk_flag(1, err = FALSE)
```

This allows developers to provide their own error message.
```{r, error = TRUE}
fun <- function(x) {
  if(!chk_flag(x, err = FALSE)) {
    stop("x really should be a flag (try as.logical())")
  }
}
fun(1)
```

In addition, `chk` is released under the [CC0](https://creativecommons.org/publicdomain/zero/1.0/) licence so developers 
can copy and paste individual functions (and tests) into their packages 
without any need for creditation which is useful if they don't want `chk` 
as a dependency.



## Inspiration

- [checkr](https://github.com/poissonconsulting/checkr/)

## Contribution

Please report any [issues](https://github.com/poissonconsulting/chk/issues).

[Pull requests](https://github.com/poissonconsulting/chk/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/poissonconsulting/chk/blob/master/CODE_OF_CONDUCT.md).
By contributing, you agree to abide by its terms
