---
title: "Using chk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using chk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`chk` is an R package for developers to check user-supplied function arguments.
It is designed to be simple, customizable and fast.

### `chk_` Functions

`chk_` functions check the properties of individual objects. 
For example `chk_flag(x)` checks whether `x` is a flag, i.e., a non-missing logical vector of length 1. 

`chk_` functions are called for their side-effects, i.e., they throw an informative error if the object fails the check.

```{r, error=TRUE}
library(chk)
y <- TRUE
chk_flag(y)
y <- NA
chk_flag(y)
```

The error messages, which follow the [tidyverse style guide](https://style.tidyverse.org/error-messages.html), are designed to allow the user to quickly identify the problem with the argument value(s) they are providing.
The errors are [rlang errors](https://rlang.r-lib.org/reference/abort.html) of subclass `'chk_error'`.

### `vld_` Functions

Each `chk_` function has a matching `vld_` function which returns a flag indicating whether the object passed the check. 
```{r, error=TRUE}
vld_flag(TRUE)
vld_flag(NA)
```
The `vld_` functions allow developers to provide their own error messages.
```{r, error=TRUE}
if(!vld_flag(NA)) stop("NA is not TRUE or FALSE!!")
```

## Using chk

The `chk_` functions are designed to be used within functions. 
Consequently, when constructing an error message they use the name of the object that they received as this is expected to be the name of the argument.
```{r, error = TRUE}
fun1 <- function(x) {
  chk_whole_number(x)
  # use x
}

fun1(1)
y <- 1.3
fun1(x = y)
```

If this is not the case, developers can provide a different name using the `x_name` argument.
```{r, error = TRUE}
x <- NA
chk_flag(x, x_name = "`zzz`")
```

**IMPORTANT NOTE**

As the `chk_` (and `vld_`) functions are not expected to be directly exposed to users they don't check any of their arguments (other than the object of interest of course!) to ensure that they are as fast as possible.

## Extending chk

The `chk_flag()` function illustrates the general structure of a `chk_` function.

``` r
chk_flag
#> function(x, x_name = NULL){
#>   if(vld_flag(x)) return(invisible())
#>   if(is.null(x_name))  x_name <- deparse_backtick_chk(substitute(x))
#>   abort_chk(x_name, " must be a flag (TRUE or FALSE)")
#> }
#> <bytecode: 0x7fe802835670>
#> <environment: namespace:chk>
```

A `chk_` function initially checks the object (using its `vld_` partner) and if the object passes the check immediately returns an invisible `NULL`.
If, and only if, the object fails the check does the `chk_` function construct and then throw an informative error message.

The `deparse_backtick_chk()` and `abort_chk()` functions are exported to make it easy for programmers to develop their own `chk_` functions.
The [chk-lgl.R](https://github.com/poissonconsulting/chk/blob/master/R/chk-lgl.R) script illustrates the general template to use when developing your own `chk_` functions.

### `abort_chk()`

The `abort_chk()` function converts multiple arguments to a string using `paste0(..., collapse = '')` and provides number sensitive `sprintf`-like types. 
By default it also capitalizes the first character and adds a missing period.
```{r, error=TRUE}
abort_chk("There %r %n problem director%y%s.", n = 1)
abort_chk("there %r %n ", "problem director%y%s", n = 2)
```

## `chk_` Families

The `chk_` functions (and their `vld_` equivalents) can be divided into the following families.
For reasons of space, the `x_name = NULL` argument is not shown.

### Check Logical Scalars

Function | Code
:- | :-
`chk_true(x)` | `is.logical(x) && length(x) == 1L && !anyNA(x) && x`
`chk_false(x)` | `is.logical(x) && length(x) == 1L && !anyNA(x) && !x`
`chk_flag(x)` | `is.logical(x) && length(x) == 1L && !anyNA(x)`
`chk_lgl(x)` | `is.logical(x) && length(x) == 1L`

### Check Scalars

Function | Code
:- | :------
`chk_scalar(x)` | `length(x) == 1L`
`chk_number(x)` | `is.numeric(x) && length(x) == 1L && !anyNA(x)`
`chk_whole_number(x)` | `vld_number(x) && (is.integer(x) || vld_true(all.equal(x, trunc(x))))`
`chk_string(x)` | `is.character(x) && length(x) == 1L && !anyNA(x)`
`chk_date(x)` | `inherits(x, "Date") && length(x) == 1L && !anyNA(x)`
`chk_datetime(x)` | `inherits(x, "POSIXct") && length(x) == 1L && !anyNA(x)`

### Check Ranges

Function  | Code
:- | :---
`chk_range(x, range = c(0, 1))` | `all(x[!is.na(x)] >= range[1] & x[!is.na(x)] <= range[2])`
`chk_lt(x, value = 0)` | `all(x[!is.na(x)] < value)`
`chk_lte(x, value = 0)` | `all(x[!is.na(x)] <= value)`
`chk_gt(x, value = 0)` | `all(x[!is.na(x)] > value)`
`chk_gte(x, value = 0)` | `all(x[!is.na(x)] >= value)`

### Check Equal

Function | Code
:-- | :-
`chk_identical(x, y)` | `identical(x, y)`
`chk_equal(x, y, tolerance = sqrt(.Machine$double.eps))` | `vld_true(all.equal(x, y, tolerance))`
`chk_equivalent(x, y, tolerance = sqrt(.Machine$double.eps))` | `vld_true(all.equal(x, y, tolerance, check.attributes = FALSE))`

### Check All

Function |  Code
:- | :---
`chk_all(x, chk_fun, ...)` | 
`chk_all_identical(x)` | `length(x) < 2L || all(vapply(x, vld_identical, TRUE, y = x[[1]]))`
`chk_all_equal(x, tolerance = sqrt(.Machine$double.eps))` | `length(x) < 2L || all(vapply(x, vld_equal, TRUE, y = x[[1]], tolerance = tolerance))`
`chk_all_equivalent(x, tolerance = sqrt(.Machine$double.eps))` | `length(x) < 2L || all(vapply(x, vld_equivalent, TRUE, y = x[[1]], tolerance = tolerance))`

### Check Set

Function | Code
:- | :---
`chk_subset(x, values)` | `all(x %in% values)`
`chk_superset(x, values)` | `all(values %in% x)`

### Check Is

Function | Code
:- | :---
`chk_s3_class(x, class)` | `!isS4(x) && inherits(x, class)`
`chk_s4_class(x, class)` | `isS4(x) && inherits(x, class)`
`chk_list(x)` | `is.list(x)`
`chk_vector(x)` | `is.vector(x)`
`chk_function(x, formals = NULL)` | `is.function(x) && (is.null(formals) || length(formals(x)) == formals)`
`chk_whole_numeric(x)` | `is.integer(x) || (is.double(x) && vld_true(all.equal(x, as.integer(x))))`

### Check NULL

Function | Code
:- | :---
`chk_null(x)` | `is.null(x)`
`chk_not_null(x)` | `!is.null(x)`

### Check ...

Function | Code
:- | :---
`chk_used(...)` | `length(list(...)) != 0L`
`chk_unused(...)` | `length(list(...)) == 0L`

### Check Files

Function |  Code
:- | :---
`chk_file(x)` | `vld_string(x) && file.exists(x) && !dir.exists(x)`
`chk_ext(x, ext)` | `vld_string(x) && vld_subset(tools::file_ext(x), ext)`
`chk_dir(x)` | `vld_string(x) && dir.exists(x)`

### Check Miscellaneous

Function | Code
:- | :--
`chk_named(x)` |  `!is.null(names(x))`
`chk_no_missing(x)` |  `!anyNA(x)`
`chk_unique(x, incomparables = FALSE)` | `!anyDuplicated(x, incomparables = incomparables)`
`chk_match(x, regexp = ".+")` | `all(grepl(regexp, x[!is.na(x)]))`
